---
title: "Big Picture"
toc: true
draft: false
format:
  html:
    resources:
      - shinylive-sw.js
filters:
  - shinylive
---
```{r load libraries}
#| results: "hide"
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(sf)
library(tidycensus)
library(here)
library(viridis)
library(dplyr)
library(MASS)
library(tidyr)

# Load crime data and convert to spatial format
data <- readRDS(here("filtered_data_2.rds"))
crime_sf <- st_as_sf(data, coords = c("LON", "LAT"), crs = 4326)

# Load neighborhood boundaries
neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp")) %>%
  st_transform(st_crs(crime_sf))

# Join crime incidents with neighborhoods
crime_with_neighborhoods <- st_join(crime_sf, neighborhoods, join = st_within)
census_api_key("c2aebe6041f0c99e41a3458ed8d0b95ee3650fa4")

# Load and transform median income data
economic_data <- get_acs(
  geography = "block group",
  variables = c(median_income = "B19013_001"),
  state = "CA",
  county = "Los Angeles",
  year = 2020,
  geometry = TRUE
) %>% 
  st_transform(st_crs(crime_with_neighborhoods))

# Join income with neighborhood-level crime data
crime_with_income <- st_join(crime_with_neighborhoods, economic_data, join = st_within)

# Summarize crime data by neighborhood
crime_summary <- crime_with_income %>%
  group_by(GEOID) %>%
  summarize(
    total_crimes = n(),
    avg_income = mean(estimate, na.rm = TRUE),
    geometry = st_union(geometry)
  ) %>%
  filter(!is.na(avg_income)) %>%  # Remove rows with missing income
  st_as_sf()



lapd_boundary <- st_read(here("CityBoundaryofLosAngeles/geo_export_416eaeda-447a-4473-b003-37e2cad181ac.shp")) %>%
  st_transform(st_crs(crime_sf))  # Align CRS



crime_within_lapd <- crime_sf %>%
  filter(rowSums(st_within(geometry, st_geometry(lapd_boundary), sparse = FALSE)) > 0)

crime_with_income_lapd <- st_join(crime_within_lapd, economic_data, join = st_within)

economic_data_lapd <- st_intersection(economic_data, lapd_boundary)


# Update the crime summary
crime_summary_lapd <- crime_with_income_lapd %>%
  group_by(GEOID) %>%
  summarize(
    total_crimes = n(),
    avg_income = mean(estimate, na.rm = TRUE),
    geometry = st_union(geometry)
  ) %>%
  filter(!is.na(avg_income)) %>%  # Exclude rows with missing income
  st_as_sf()

crime_summary_race <- crime_with_income %>%
  group_by(GEOID, Descent_Description) %>%
  summarize(
    total_crimes = n(),  # Count crimes for each GEOID and Descent_Description group
    avg_median_income = mean(estimate, na.rm = TRUE)
  ) %>%
  pivot_wider(
    names_from = Descent_Description,
    values_from = total_crimes,
    values_fill = 0  # Fill missing values with 0 for crimes
  ) %>%
  rowwise() %>%
  mutate(
    total_crimes = sum(c_across(-c(GEOID, avg_median_income, geometry)), na.rm = TRUE)  # Sum only the crime columns
  ) %>%
  ungroup()

crime_summary_race <- crime_summary_race %>%
  rename_with(~ make.names(.), everything())

crime_summary_race_model <- crime_summary_race %>%
  filter(complete.cases(
    avg_median_income, Hispanic.Latin.Mexican, White, Black, Other, Other.Asian,
    Unknown, Japanese, Korean, Filipino, Chinese, Hawaiian, Asian.Indian,
    American.Indian.Alaskan.Native, Cambodian, Vietnamese, Laotian
  ))


bivar_nb_model <- glm.nb(total_crimes ~ avg_median_income + Hispanic.Latin.Mexican + White + Black + Other + 
    Other.Asian + Unknown + Japanese + Korean + Filipino + Chinese + Hawaiian + 
    Asian.Indian + American.Indian.Alaskan.Native + Cambodian + Vietnamese + Laotian, data = crime_summary_race)

# Add predicted values to the filtered data
crime_summary_race_model <- crime_summary_race_model %>%
  mutate(predicted_crimes = predict(bivar_nb_model, type = "response"))

crime_race_long <- crime_summary_race_model %>%
  pivot_longer(
    cols = c(Hispanic.Latin.Mexican, White, Black, Other, Other.Asian, Unknown, 
             Japanese, Korean, Filipino, Chinese, Hawaiian, Asian.Indian, 
             American.Indian.Alaskan.Native, Cambodian, Vietnamese, Laotian),
    names_to = "race",
    values_to = "crime_count"
  )


crime_race_long <- crime_summary_race %>%
  pivot_longer(
    cols = c(Hispanic.Latin.Mexican, White, Black, Other, Other.Asian, Unknown, 
             Japanese, Korean, Filipino, Chinese, Hawaiian, Asian.Indian, 
             American.Indian.Alaskan.Native, Cambodian, Vietnamese, Laotian),
    names_to = "race",
    values_to = "crime_count"
  ) %>%
  filter(!is.na(crime_count) & crime_count > 0) %>%
  mutate(avg_median_income = avg_median_income / 1000)



# Fit Poisson regression with pivoted data
poisson_model <- glm(
  crime_count ~ avg_median_income + race,
  family = poisson(link = "log"),
  data = crime_race_long
)
summary(poisson_model)

# Fit Negative Binomial regression with pivoted data
crime_race_long <- crime_race_long %>%
  filter(complete.cases(avg_median_income, race, crime_count))

nb_model <- glm.nb(
  crime_count ~ avg_median_income * race,
  data = crime_race_long
)

crime_race_long <- crime_race_long %>%
  mutate(predicted_crimes = predict(nb_model, newdata = crime_race_long, type = "response"))


lapd_boundary <- st_read(here("CityBoundaryofLosAngeles/geo_export_416eaeda-447a-4473-b003-37e2cad181ac.shp")) %>%
  st_make_valid() %>%                # Ensure valid geometry
  st_transform(crs = 4326)           # Align CRS (use WGS84 as default)

# Load neighborhoods
neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp")) %>%
  st_make_valid() %>%
  st_transform(crs = st_crs(lapd_boundary))  # Match CRS with LAPD boundary


lapd_boundary <- st_read(here("CityBoundaryofLosAngeles/geo_export_416eaeda-447a-4473-b003-37e2cad181ac.shp")) %>%
  st_make_valid() %>%                # Ensure valid geometry
  st_transform(crs = 4326)           # Align CRS (use WGS84 as default)

# Load neighborhoods
neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp")) %>%
  st_make_valid() %>%
  st_transform(crs = st_crs(lapd_boundary))  # Match CRS with LAPD boundary


population_data <- get_acs(
  geography = "block group",
  variables = c(total_population = "B01003_001"),
  state = "CA",
  county = "Los Angeles",
  year = 2020,
  geometry = TRUE
)

neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp")) %>%
  st_make_valid() %>%
  st_transform(st_crs(population_data))


population_with_neighborhoods <- st_join(population_data, neighborhoods, join = st_within)


# Summarize total population by neighborhood
neighborhood_population <- population_with_neighborhoods %>%
  group_by(name) %>%  # Assuming 'name' is the neighborhood name column
  summarize(total_population = sum(estimate, na.rm = TRUE)) %>%
  filter(!is.na(total_population))  # Exclude rows with missing population


areaData <- readRDS(here("crime_counts_by_area_with_coords.rds"))

# Convert crime data into a spatial object (assuming coordinates are in 'avg_longitude' and 'avg_latitude')
crime_data_sf <- st_as_sf(areaData, coords = c("avg_longitude", "avg_latitude"), crs = 4326)

# Ensure that the neighborhood population data and neighborhoods are in the same CRS
neighborhood_population <- st_transform(neighborhood_population, st_crs(crime_data_sf))
neighborhoods <- st_transform(neighborhoods, st_crs(crime_data_sf))

# Ensure both datasets have valid geometries
crime_data_sf <- st_make_valid(crime_data_sf)
neighborhood_population <- st_make_valid(neighborhood_population)

# Check the CRS of both datasets to ensure they match
print(st_crs(crime_data_sf))  # CRS of crime data
print(st_crs(neighborhood_population))  # CRS of neighborhood population

# If necessary, transform the neighborhood population data to the same CRS as crime data
# This step should be redundant if CRS checks show they already match
neighborhood_population <- st_transform(neighborhood_population, st_crs(crime_data_sf))

# Perform the spatial join to associate crime data with the neighborhood polygons
crime_population_with_data <- st_join(crime_data_sf, neighborhood_population, join = st_within)

# Now calculate the crime-to-population ratio for each record
crime_population_summary <- crime_population_with_data %>%
  mutate(crime_to_population_ratio = crime_count / total_population)  # Divide total crime by total population




```

```{shinylive-r}
#| results: "hide"
#| echo: false
#| standalone: true
#| viewerHeight: 640
library(tidyverse)
library(sf)
library(tidycensus)
library(here)
library(viridis)
library(dplyr)
library(MASS)
library(tidyr)
library(shiny)

options("readr.edition" = 1)
crime_with_income_lapd <- read_rds("https://sussmanbu.github.io/ma-4615-fa24-final-project-group-4/shiny_dataset/crime_with_income_lapd.rds")
economic_data_lapd<-read_rds("https://sussmanbu.github.io/ma-4615-fa24-final-project-group-4/shiny_dataset/economic_data_lapd.rds")

ui <- fluidPage(
  titlePanel("Crime and Economic Data in LAPD Jurisdiction"),
  sidebarLayout(
    sidebarPanel(
      sliderInput(
        inputId = "year",
        label = "Select Year:",
        min = 2020, 
        max = 2024, 
        value = 2020,
        step = 1
      )
    ),
    mainPanel(
      plotOutput("crimePlot")
    )
  )
)

server <- function(input, output, session) {
  output$crimePlot <- renderPlot({
    crime_filtered <- crime_with_income_lapd %>%
      filter(year == input$year)
    
    ggplot() +
      geom_sf(data = economic_data_lapd, aes(fill = estimate), color = NA) +  
      scale_fill_viridis_c(option = "magma", na.value = "grey50") +
      geom_sf(data = crime_filtered, color = "black", shape = 4, size = 0.5, alpha = 0.7) +
      labs(
        title = paste("Median Income and", input$year, "Crime Data within LAPD Jurisdiction"),
        fill = "Median Income"
      ) +
      theme_minimal() +
      coord_sf(xlim = c(-118.7, -118), ylim = c(33.7, 34.5),expand = FALSE)
  })
}

shinyApp(ui = ui, server = server)
```


## Beyond the Poverty Myth: Unraveling Los Angeles' Crime Landscape


When you think about crime, you might picture dangerous streets in low-income neighborhoods and pristine, secure communities in wealthier areas. Normally, you would avoid that one sketchy neighborhood when it is late at night. It’s a common narrative, repeated so often that it feels like fact: crime is a symptom of poverty, and wealth provides a safety net against it. But is this common perception really the truth? 

We attempt to take a closer look at the relationship between income levels and crime rates in Los Angeles, aiming to uncover patterns that challenge conventional wisdom about the link between poverty and crime. Using the LAPD crime reports and Census data from the past 4 years, we set out to explore the connection between income levels and crime rates across different neighborhoods in Los Angeles. The results might surprise you. While crime does concentrate in economically disadvantaged areas, the relationship between median income and crime is far from straightforward—and some of our findings challenge these well-worn assumptions. The stereotype that low median areas experience higher crime rates is overly simplistic. While violent crimes show some correlation with lower-income neighborhoods, broader crime patterns reveal that income alone does not explain crime rates. Our analysis uncovers complexities and outliers that challenge conventional beliefs. 

### The Story Behind the Data
When you think about crime, certain images likely come to mind—run-down neighborhoods with boarded-up windows or wealthy enclaves guarded by private security. But how do crime rates really distribute across the city, and what role does income play?

To answer these questions, we dove into LAPD crime reports and Census data from the past four years, breaking down crime trends by income, neighborhood population, and crime type. We looked beyond total crime counts to analyze violent crimes, population-adjusted metrics, and outliers that defy expectations.

Here’s what we examined:

* Which neighborhoods experience the highest total crime rates?
* How does violent crime correlate with median income?
* Which areas have the highest crime-to-population ratios?

Through this analysis, we paint a clearer picture of crime patterns in Los Angeles, revealing where the traditional narratives hold true—and where they fall apart.

### Visualizing Crime Trends

```{r heatmap for crime count}
#| echo: false
#| warning: false

neighborhoods_lapd <- st_intersection(neighborhoods, lapd_boundary)


neighborhoods_lapd <- st_buffer(neighborhoods, dist = 0) %>%
  st_intersection(st_buffer(lapd_boundary, dist = 0))


crime_sf <- st_as_sf(data, coords = c("LON", "LAT"), crs = 4326)  # Crime data as sf object

# Filter crimes to those within LAPD boundary
crimes_in_lapd <- st_filter(crime_sf, lapd_boundary)

neighborhoods_lapd <- st_make_valid(neighborhoods_lapd)
crimes_in_lapd <- st_make_valid(crimes_in_lapd)

# Join crimes to neighborhoods
crimes_by_neighborhood <- st_join(crimes_in_lapd, neighborhoods_lapd, join = st_within)


# Ensure crimes_by_neighborhood includes counts and neighborhood polygons
crime_summary <- neighborhoods_lapd %>%
  left_join(
    crimes_by_neighborhood %>%
      st_drop_geometry() %>%  # Drop point geometry to focus on attributes
      group_by(name) %>%      # Replace "NAME" with neighborhood column in neighborhoods_lapd
      summarize(total_crimes = n()),
    by = "name"               # Join by neighborhood name or another matching column
  )


ggplot() +
  geom_sf(data = neighborhoods_lapd, fill = NA, color = "black") +  # Neighborhood outlines
  geom_sf(data = crime_summary, aes(fill = total_crimes), color = NA) +
  scale_fill_distiller(
    palette = "YlOrRd", 
    direction = 1, 
    na.value = "grey50",
    name = "Total Crimes"
  ) +
  labs(
    title = "Heatmap of Crimes by Neighborhood (LAPD Jurisdiction)",
    fill = "Total Crimes"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  )

```
Using LAPD data, we mapped total crime counts across Los Angeles neighborhoods. The heatmap shows darker areas for lower crime counts and lighter areas for higher crime activity. Darker areas indicate lower crime counts, while lighter areas signify higher crime activity. At a glance, neighborhoods in Central LA and parts of West LA emerge as hotspots for criminal activity.

However, this heatmap only shows the total crime counts geographically, without considering the underlying economic conditions of each neighborhood. By incorporating median income into the data and visualizations, we can better understand the relationship between crime and median income.

### Median Income vs. Total Crimes

```{r relationship between crime and income}
#| echo: false
crime_with_income <- st_join(crime_with_neighborhoods, economic_data, join = st_within)
median_income_by_neighborhood <- crime_with_income %>%
  group_by(AREA.NAME) %>%
  summarize(
    median_income = median(estimate, na.rm = TRUE),
    total_crimes = n(),  # Optional: Include total crimes
    geometry = st_union(geometry)
  ) %>%
  filter(!is.na(median_income)) %>% 
  st_as_sf()

# Create new columns for scaled axes
median_income_by_neighborhood <- median_income_by_neighborhood %>%
  mutate(
    median_income_scaled = median_income / 1000,
    total_crimes_scaled = total_crimes
  )

ggplot(median_income_by_neighborhood, aes(x = median_income_scaled, y = total_crimes_scaled)) +
  geom_point(color = "blue", alpha = 0.7, size = 4) +
  labs(
    title = "Relationship Between Median Income and Total Crimes",
    x = "Median Income (USD in thousands)",
    y = "Total Crimes",
    caption = "Source: ACS 2020 and Crime Data"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",  # Remove legend as color is no longer mapped
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```
A scatterplot illustrates the relationship between median income and total crime rates for each neighborhood. Each point represents a neighborhood, with median income on the x-axis and total crimes on the y-axis. Points are color-coded, ranging from dark purple (lower income) to bright yellow (higher income).

Interestingly, the scatterplot reveals no consistent correlation between income and total crime. While some low-income neighborhoods see higher crime rates, several affluent areas also report significant crime activity. This challenges the simplistic view that wealth equals safety, suggesting other factors—like population density and social dynamics—play a role.


### Delving Into Violent Crime

```{r violent crime in each neighborhood}
#| echo: false

crime_with_income <- crime_with_income %>%
  mutate(
    crime_type = case_when(
      str_detect(Crm.Cd.Desc, regex("Burglary|Theft|Shoplifting|Robbery|Vandalism|Stolen", ignore_case = TRUE)) ~ "Property Crime",
      str_detect(Crm.Cd.Desc, "ASSAULT|ROBBERY|HOMICIDE|KIDNAPPING|SEXUAL ASSAULT") ~ "Violent Crime",
      TRUE ~ "Other"
    )
  )
property_crime_summary <- crime_with_income %>%
  filter(crime_type == "Violent Crime") %>%  # Filter for property crimes
  group_by(AREA.NAME) %>%  # Group by neighborhood
  summarize(
    median_income = median(estimate, na.rm = TRUE),  # Median income per neighborhood
    property_crimes = n()  # Count the number of property crimes
  ) %>%
  arrange(median_income) %>%  # Order neighborhoods by median income
  mutate(AREA.NAME = factor(AREA.NAME, levels = AREA.NAME))  # Convert to factor to preserve order

# Plot the data
ggplot(property_crime_summary, aes(x = AREA.NAME, y = property_crimes)) +
  geom_bar(stat = "identity", aes(fill = median_income), show.legend = TRUE) +
  scale_fill_viridis_c(option = "plasma", name = "Median Income") +
  labs(
    title = "Violent
    Crimes vs Median Income by Neighborhood",
    x = "Neighborhood (Ordered by Median Income)",
    y = "Number of Violent Crimes",
    caption = "Source: ACS 2020 and Crime Data"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    plot.title = element_text(hjust = 0.5)
  )

```

When we shift focus to violent crime, a more distinct pattern emerges. Bar charts sorting neighborhoods by median income reveal a negative correlation: lower-income areas like Newton and Rampart report significantly higher violent crime rates, while wealthier neighborhoods like West LA and Devonshire see fewer incidents.

There is a glaring outlier of Central LA facing high crime counts despite being on the higher median income relative to the other Los Angeles Neighborhood which doesn't fit with the correlation between median income and violent crime. However, this only delves into a certain category of crime. Crime in Los Angeles is a multifaceted issue. Our analysis challenges the simplistic narrative linking poverty and crime, showing that even wealthy neighborhoods can face significant crime challenges. 

While focusing on violent crimes reveals a clear negative correlation with median income, examining crime relative to neighborhood populations adds another layer of complexity. Being in more dense-populated neighborhoods may contribute to more crime which scaling the crime count in proportion to population give a more accurate look into the relationship with median income and crime.

### Crime-to-Population Ratios

```{r Ratio Graph}
#| echo: false

library(ggplot2)
library(dplyr)

# Filter for 2024 data, remove NA values, and calculate crime-to-population ratio
crime_population_2024 <- crime_population_summary %>%
  filter(year == 2024) %>%
  mutate(crime_to_population_ratio = ifelse(total_population > 0, crime_count / total_population, NA)) %>%
  filter(!is.na(crime_to_population_ratio) & !is.na(`AREA NAME`))  # Remove rows with NA in crime ratio or area name

# Plot the crime-to-population ratio per area for 2024
ggplot(crime_population_2024, aes(x = reorder(`AREA NAME`, crime_to_population_ratio), y = crime_to_population_ratio)) +
  geom_bar(stat = "identity", fill = "blue") +  # Set the bar color to blue
  coord_flip() +  # Flip the axes
  labs(
    title = "Crime-to-Population Ratio per Area (2024)",
    x = "Neighborhood",
    y = "Crime-to-Population Ratio"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))  # Adjust text size for readability

```
To account for population differences, we analyzed crime-to-population ratios across neighborhoods. Areas like 77th Street, Southwest, and Southeast show disproportionately high ratios, meaning residents in these neighborhoods face crime at significantly higher rates. Unsurprisingly, these neighborhoods tend to fall on the lower end of the median income spectrum.

### Key Takeaways

The relationship between median income and crime in Los Angeles is far from straightforward. While there is a noticeable negative correlation—lower median income neighborhoods often experience higher crime rates—this relationship becomes less clear when accounting for factors like population density and proportional impacts. Population size, density, and local factors significantly shape crime distribution across neighborhoods. A simple link between poverty and crime oversimplifies a complex issue. Addressing crime in Los Angeles requires considering economic conditions, community dynamics, and residents' lived experiences.

```{shinylive-r}
#| results: "hide"
#| echo: false
#| standalone: true
#| viewerHeight: 640
library(shiny)
library(dplyr)
library(ggplot2)
options("readr.edition" = 1)
areaData=read_rds("https://sussmanbu.github.io/ma-4615-fa24-final-project-group-4/shiny_dataset/areaData.rds")
colnames(areaData)[colnames(areaData) == "AREA NAME"] <- "AREA_NAME"

# Ensure year is numeric:
areaData$year <- as.numeric(areaData$year)

ui <- fluidPage(
  titlePanel("Los Angeles Area Crime Dashboard"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput(
        inputId = "selected_area",
        label = "Select an Area:",
        choices = unique(areaData$AREA_NAME),
        selected = unique(areaData$AREA_NAME)[1] # default to first area
      )
    ),
    mainPanel(
      h2("Area Summary"),
      verbatimTextOutput("summary_stats"),
      
      h3("Total Crime Over Time"),
      plotOutput("crime_trend_plot")
    )
  )
)

server <- function(input, output, session) {
  
  # Filter data based on selected area
  selected_data <- reactive({
    req(input$selected_area)
    areaData %>% 
      filter(AREA_NAME == input$selected_area)
  })
  
  # Summary text
  output$summary_stats <- renderPrint({
    d <- selected_data()
    # Calculate some basic summary information
    yrs <- range(d$year)
    total_min <- min(d$crime_count, na.rm = TRUE)
    total_max <- max(d$crime_count, na.rm = TRUE)
    avg_lat <- mean(d$avg_latitude, na.rm = TRUE)
    avg_lon <- mean(d$avg_longitude, na.rm = TRUE)
    
    cat("Selected Area:", input$selected_area, "\n",
        "Year Range:", paste0(yrs[1], " - ", yrs[2]), "\n",
        "Min Crime Count:", total_min, "\n",
        "Max Crime Count:", total_max, "\n",
        "Average Latitude:", round(avg_lat, 5), "\n",
        "Average Longitude:", round(avg_lon, 5), "\n",
        "Number of Records:", nrow(d), "\n")
  })
  
  # Line plot of total crime over time
  output$crime_trend_plot <- renderPlot({
    d <- selected_data()
    ggplot(d, aes(x = year, y = crime_count)) +
      geom_line(color = "steelblue", size = 1) +
      geom_point(color = "steelblue", size = 2) +
      labs(title = paste("Crime Trends in", input$selected_area),
           x = "Year",
           y = "Total Crime") +
      theme_minimal()
  })
}

shinyApp(ui = ui, server = server)

```

