---
title: Data
description: We describe the sources of our data and the cleaning process.
toc: true
draft: false
---

# Analysis and Cleaning of LAPD Crime Data (2020–Present)

# Data Documentation

## 1. Data Overview

The dataset contains reported crimes in Los Angeles from 2020 onward, sourced from the **LAPD**.

- **Source**: [LAPD Crime Data (2020–Present)](https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/data_preview)
- **Attribution**: Derived from original crime reports typed on paper.
- **Purpose**:
  - Track crime patterns.
  - Allocate public safety resources.
  - Promote transparency and public accountability.
- **Challenges**:
  - Large dataset size (~987K rows, 28 variables).
  - Cleaning required to address duplicates, incomplete data, and usability for spatial/temporal analysis.

---

## 2. Variables in the Dataset

The dataset comprises 28 columns, described below:

| **Variable**        | **Description**                                                                                   | **Data Type**         |
|----------------------|---------------------------------------------------------------------------------------------------|------------------------|
| DR_NO               | Division of Records Number: Unique file identifier (YY+Area ID+5 digits).                        | Text                  |
| Date Rptd           | Date crime was reported (MM/DD/YYYY).                                                            | Timestamp             |
| DATE OCC            | Date crime occurred (MM/DD/YYYY).                                                                | Timestamp             |
| TIME OCC            | Time crime occurred (24-hour format).                                                            | Text                  |
| AREA                | LAPD’s 21 geographic areas (1–21).                                                               | Text                  |
| AREA NAME           | Name corresponding to the geographic area (e.g., 77th Street).                                   | Text                  |
| Rpt Dist No         | Sub-area code within a geographic area.                                                          | Text                  |
| Part 1-2            | Classification of crime as Part I or Part II.                                                    | Number                |
| Crm Cd              | Primary crime code.                                                                               | Text                  |
| Crm Cd Desc         | Description of the primary crime.                                                                | Text                  |
| Mocodes             | Suspect activities during the crime.                                                             | Text                  |
| Vict Age            | Age of the victim.                                                                               | Number                |
| Vict Sex            | Victim’s gender (M: Male, F: Female, X: Unknown).                                                | Text                  |
| Vict Descent        | Ethnic descent (e.g., H: Hispanic, W: White).                                                    | Text                  |
| Premis Cd           | Code for the type of location where the crime occurred.                                          | Number                |
| Premis Desc         | Description of the premise code.                                                                 | Text                  |
| Weapon Used Cd      | Code for the weapon used.                                                                        | Text                  |
| Weapon Desc         | Description of the weapon code.                                                                  | Text                  |
| Status              | Status of the case (e.g., IC: Incomplete).                                                       | Text                  |
| Status Desc         | Description of the status code.                                                                  | Text                  |
| Crm Cd 1–4          | Primary and additional crime codes.                                                              | Text                  |
| LOCATION            | Approximate address of the crime (anonymized).                                                  | Text                  |
| Cross Street        | Cross street near the crime location.                                                            | Text                  |
| LAT                 | Latitude of the crime location.                                                                  | Number                |
| LON                 | Longitude of the crime location.                                                                 | Number                |

---

## 3. Cleaning and Sampling

### Initial Dataset Loading
The dataset was loaded as a CSV file for easier manipulation. Here’s the R code for loading it:

```{r}
#| results: hide
#| warning: false
library(here)
library(readr)
library(dplyr)
library(tidyverse)
data <- read_csv(here("your_file.csv"))
```
Then we start the data cleaning process. 

To reduce the dataset size (because it was too big previously) and balance representation across years, we:\

- Extracted the year from the DATE OCC column.\
- Filtered data for years 2020–2024.\
- Sampled up to 2,500 rows per year for efficiency.\

The following r chunk is the exact code we use by following the previous steps.
```{r}

# Convert the date column to year format
data$year <- format(as.Date(data$DATE.OCC, format = "%m/%d/%Y"), "%Y")
# Convert year to numeric
data$year <- as.numeric(data$year)  
# Filter for the years 2020–2024
filtered_years <- data %>%
  filter(year %in% 2020:2024)
# Split by year and sample up to 2,500 rows per year
sampled_data <- filtered_years %>%
  group_split(year) %>%
  map_df(~ slice_sample(.x, n = min(2500, nrow(.x))))
```

The RDS file after first time cleaning can look like this (some columns filtered out for readability):
```{r}
#| echo: False 
#| results: hide
datPage = readRDS(here("filtered_data_copy.rds"))
```
## Updated Dataset with Years Spread Across Multiple Years

| **DR_NO**   | **DATE.OCC**        | **Year** | **AREA.NAME** | **Crm.Cd** | **Crm.Cd.Desc**                        | **Premis.Desc**           | **LAT**     | **LON**      |
|-------------|---------------------|----------|---------------|------------|----------------------------------------|---------------------------|-------------|--------------|
| 200604124   | 01/03/2020          | 2020     | Hollywood     | 310        | BURGLARY                               | RESTAURANT/FAST FOOD      | 34.1000     | -118.3113    |
| 200218223   | 12/05/2019          | 2019     | Rampart       | 510        | VEHICLE - STOLEN                       | STREET                    | 34.0707     | -118.2693    |
| 210105387   | 02/02/2021          | 2021     | Central       | 649        | DOCUMENT FORGERY / STOLEN FELONY       | BANK                      | 34.0493     | -118.2582    |
| 201314683   | 07/29/2020          | 2020     | Newton        | 740        | VANDALISM - FELONY                     | VEHICLE, PASSENGER/TRUCK  | 33.9830     | -118.2783    |
| 201610483   | 06/28/2022          | 2022     | Foothill      | 354        | THEFT OF IDENTITY                      | SINGLE FAMILY DWELLING    | 34.2103     | -118.3747    |
| 200905047   | 01/25/2019          | 2019     | Van Nuys      | 745        | VANDALISM - MISDEMEANOR ($399 OR UNDER)| VEHICLE STORAGE LOT       | 34.1961     | -118.4487    |
| 200111368   | 05/05/2021          | 2021     | Central       | 940        | EXTORTION                              | MULTI-UNIT DWELLING       | 34.0433     | -118.2377    |
| 201508876   | 04/16/2022          | 2022     | N Hollywood   | 310        | BURGLARY                               | LAUNDROMAT                | 34.1649     | -118.3615    |

# Census Data Utilized in Crime Analysis
To enhance the depth of our crime analysis, we incorporated U.S. Census data to provide socioeconomic context for various geographic areas. The census data offers crucial insights into factors such as population size, median income, and racial composition, which are essential for understanding potential correlations between crime rates and socioeconomic variables.However, we wanted to look just at median income and take the average of the median income for an area. 

For this analysis, the census data was matched with specific crime locations, enabling us to explore whether crime rates differ significantly based on income levels. Median income was a particularly important variable, used to examine potential relationships between socioeconomic status and crime prevalence. Median income was used instead of average income due to high income areas which might have extreme outliers with certain individuals making a large amount of money. 

We ensured the census data was updated every decade, so this comes from 2020 but is mapped to the corresponding regions where crimes occurred. This allowed for a comprehensive analysis that moves beyond raw crime figures and considers underlying socioeconomic conditions, facilitating a more nuanced interpretation of trends and patterns.

By integrating census data, we aimed to uncover deeper insights, such as whether economically disadvantaged areas experience higher crime rates.

# Crimes in Neighborhoods:

Instead of just looking at the 12,500 crimes that we initially filtered for the sake of files being too large, it was suggested that geographic data would be filtered already as there were a lot of crimes which were being left out which could lead to bias,. So, we went back to the dataset and got all the crimes that had a latitude and longitude in the LA City era. Here is the code for that:

```{data}
#| echo: False
#| warning: false
#| results: "hide"
#| message: false
library(dplyr)
library(readr)

# Read the CSV file
data <- read_csv("Crime_Data_from_2020_to_Present_20241018.csv")

# Convert 'DATE OCC' to Date and extract the year
data$year <- as.numeric(format(as.Date(data$`DATE OCC`, format = "%m/%d/%Y"), "%Y"))

# Filter out rows with missing AREA NAME, LAT, or LON values
filtered_data <- data %>%
  filter(!is.na(`AREA NAME`) & `AREA NAME` != "" & !is.na(LAT) & !is.na(LON))

# Select relevant columns: AREA NAME, year, LAT, LON
selected_data <- filtered_data %>%
  select(`AREA NAME`, year, LAT, LON)

# Count the number of crimes by AREA NAME and year
crime_counts_by_area <- selected_data %>%
  group_by(`AREA NAME`, year) %>%
  summarise(
    crime_count = n(),
    avg_latitude = mean(LAT, na.rm = TRUE),
    avg_longitude = mean(LON, na.rm = TRUE),
    .groups = "drop"
  )

# View the summarized crime counts with coordinates
print(crime_counts_by_area)

# Save the summarized data as an RDS file
saveRDS(crime_counts_by_area, "crime_counts_by_area_with_coords.rds")

```

Here we got a new RDS File and we got the crime amounts in the different regions for a specific year from 2020-2024. Since not all years had data, the heatmap you see on the big picture page is of the year 2024 which had data from all regions of LA. 

Here is what the data could look like after cleaning. 

| **AREA NAME** | **Year** | **Crime Count** | **Avg Latitude** | **Avg Longitude** |
|---------------|----------|-----------------|-----------------|------------------|
| 77th Street   | 2024     | 13343           | 33.86060        | -117.8887        |
| Central       | 2024     | 11600           | 33.84789        | -117.5570        |
| Devonshire    | 2024     | 7982            | 34.06148        | -117.8855        |
| Foothill      | 2024     | 7106            | 34.13832        | -117.9937        |
| Harbor        | 2024     | 8874            | 33.67433        | -117.9386        |
| Hollenbeck    | 2024     | 7805            | 33.88906        | -117.6273        |
| Hollywood     | 2024     | 10171           | 33.87138        | -117.5391        |
| Mission       | 2024     | 8480            | 34.05963        | -117.7660        |
| N Hollywood   | 2024     | 10167           | 34.05776        | -117.9891        |
| Newton        | 2024     | 9994            | 33.87209        | -117.7877        |
| Northeast     | 2024     | 8452            | 34.02285        | -117.9502        |
| Olympic       | 2024     | 9639            | 33.91484        | -117.7972        |
| Pacific       | 2024     | 11575           | 33.79474        | -117.7623        |
| Rampart       | 2024     | 9025            | 33.90471        | -117.7237        |
| Southeast     | 2024     | 10847           | 33.77965        | -117.7112        |
| Southwest     | 2024     | 11178           | 33.91014        | -117.9367        |
| Topanga       | 2024     | 8106            | 34.13290        | -118.3959        |
| Van Nuys      | 2024     | 8763            | 34.11259        | -118.2174        |
| West LA       | 2024     | 9309            | 34.01168        | -118.3017        |
| West Valley   | 2024     | 8091            | 34.13708        | -118.3457        |
| Wilshire      | 2024     | 9291            | 33.89677        | -117.7792        |




