---
title: "blog post 6"
author: "Sean Fung, Zihao Guo, Yawen Zhang"
date: "2024-11-13"
date-modified: "2024-11-13"
draft: FALSE
---

```{r load libraries}
#| results: "hide"
#| echo: false
library(tidyverse)
library(sf)
library(tidycensus)
library(here)
library(viridis)

# Load crime data and convert to spatial format
data <- readRDS(here("filtered_data_2.rds"))
crime_sf <- st_as_sf(data, coords = c("LON", "LAT"), crs = 4326)

# Load neighborhood boundaries
neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp")) %>%
  st_transform(st_crs(crime_sf))

# Join crime incidents with neighborhoods
crime_with_neighborhoods <- st_join(crime_sf, neighborhoods, join = st_within)
census_api_key("c2aebe6041f0c99e41a3458ed8d0b95ee3650fa4")

# Load and transform median income data
economic_data <- get_acs(
  geography = "block group",
  variables = c(median_income = "B19013_001"),
  state = "CA",
  county = "Los Angeles",
  year = 2020,
  geometry = TRUE
) %>% 
  st_transform(st_crs(crime_with_neighborhoods))

# Join income with neighborhood-level crime data
crime_with_income <- st_join(crime_with_neighborhoods, economic_data, join = st_within)

# Summarize crime data by neighborhood
crime_summary <- crime_with_income %>%
  group_by(GEOID) %>%
  summarize(
    total_crimes = n(),
    avg_income = mean(estimate, na.rm = TRUE),
    geometry = st_union(geometry)
  ) %>%
  filter(!is.na(avg_income)) %>%  # Remove rows with missing income
  st_as_sf()
```
# Crime V Median Income
```{r Crime V Median Income Plot}
#| echo: false
# Plot: Crime vs Median Income Scatter Plot
ggplot(crime_summary, aes(x = avg_income, y = total_crimes)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Crime vs Median Income", x = "Median Income", y = "Total Crimes")
```

# Economic Heatmap (Without Crime Data)

```{r Income heat map (see bp5)}
#| echo: false
ggplot() +
  geom_sf(data = economic_data, aes(fill = estimate), color = NA) +  # Median income layer
  scale_fill_viridis_c(option = "magma", na.value = "grey50") +
  scale_color_viridis_c(option = "inferno", na.value = "grey50") +
  labs(
    title = "Median Income and Crime Data by Neighborhood in Los Angeles",
    fill = "Median Income",
    color = "Average Income"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  coord_sf(
    xlim = c(-119, -117),  # Adjust these values to focus on the Los Angeles area
    ylim = c(33, 35)       # Adjust these values to focus on the Los Angeles area
  )
```

# Plot: Combined Map with Income and Crime Data -- Black Dots (Crime Locations) on Economic Data Heatmap
```{r Heatmap with crime data}
#| echo: false
ggplot() +
  # Economic data as a heatmap
  geom_sf(data = economic_data, aes(fill = estimate), color = NA, alpha = 0.8) +
  # Black dots for crime data
  geom_sf(data = crime_sf, color = "black", size = 0.5, alpha = 0.7) +
  scale_fill_viridis_c(option = "magma", na.value = "grey50") +
  labs(
    title = "Crime Incidents on Median Income Heatmap",
    fill = "Median Income"
  ) +
  coord_sf(
    xlim = c(-119, -117),  # Focus on Los Angeles area
    ylim = c(33, 35)       # Focus on Los Angeles area
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

# Linear Model
```{r Linear model}
#| echo: false
# Linear Model: Relationship Between Crimes and Income
model <- lm(total_crimes ~ avg_income, data = crime_summary)
summary(model)

```
Based on the above models, we need to be looking at the certain areas with crime data as the LAPD seems to be concentrated in the multiple districts in the Southwest. Try to adjust for bias. 

## Crime in each neighborhood
```{r}
# Define broader crime categories based on `Crm.Cd.Desc`
crime_with_neighborhoods <- crime_with_neighborhoods %>%
  mutate(Crime_Category = case_when(
    str_detect(Crm.Cd.Desc, "BURGLARY|THEFT|VANDALISM|EMBEZZLEMENT|FORGERY|SHOPLIFTING") ~ "Property Crime",
    str_detect(Crm.Cd.Desc, "ASSAULT|ROBBERY|HOMICIDE|KIDNAPPING|SEXUAL ASSAULT") ~ "Violent Crime",
    str_detect(Crm.Cd.Desc, "DRUG") ~ "Drug Offense",
    str_detect(Crm.Cd.Desc, "TRAFFIC|DRIVING|DUI|FAILURE TO YIELD") ~ "Traffic Violation",
    TRUE ~ "Other"
  ))
crime_summary <- crime_with_neighborhoods %>%
  group_by(AREA.NAME, Crime_Category) %>%
  summarize(total_crimes = n(), .groups = "drop")
ggplot(crime_summary, aes(x = reorder(AREA.NAME, total_crimes), y = total_crimes, fill = Crime_Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Crime Categories by Neighborhood",
    x = "Neighborhood",
    y = "Total Crimes",
    fill = "Crime Category"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
```{r}
ggplot(crime_summary, aes(x = reorder(AREA.NAME, total_crimes), y = total_crimes, fill = Crime_Category)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    title = "Proportion of Crime Categories by Neighborhood",
    x = "Neighborhood",
    y = "Proportion",
    fill = "Crime Category"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
These graphs tell a compelling story about crime patterns in neighborhoods. Property crimes are a widespread issue, while violent crimes are relatively localized. Hollywood and similar areas show distinctive crime patterns, suggesting localized factors influencing crime dynamics.The 77th Street neighborhood stands out in both graphs due to the almost equal distribution of Property Crime and Other Crime categories. This is unusual compared to most neighborhoods where Property Crime significantly dominates.
```{r}
library(dplyr)
library(ggplot2)
set.seed(123) 

kmeans_data <- crime_with_income %>%
  group_by(AREA.NAME) %>%
  summarize(
    avg_income = mean(estimate, na.rm = TRUE),  
    total_crimes = n()  
  ) %>%
  na.omit() 
kmeans_data_numeric <- kmeans_data %>%
  st_drop_geometry() %>%  
  select(avg_income, total_crimes)  
kmeans_result <- kmeans(kmeans_data_numeric, centers = 3)
kmeans_data <- kmeans_data %>%
  mutate(cluster = factor(kmeans_result$cluster))


ggplot(kmeans_data, aes(x = avg_income, y = total_crimes, color = cluster)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "clustering neighborhoods by income and crime",
    x = "Median Income",
    y = "Total Crimes",
    color = "Cluster"
  ) +
  theme_minimal()


```
cluster 1: moderate crime counts and moderate income levels
cluster 2: higher income levels and lower crime counts
cluster 3: high crime counts and lower income levels 
```{r}
cluster_centers <- as.data.frame(kmeans_result$centers)

ggplot(kmeans_data, aes(x = avg_income, y = total_crimes, color = cluster)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_point(data = cluster_centers, aes(x = avg_income, y = total_crimes), size = 5, shape = 17, color = "black") + 
  labs(
    title = "clustering neighborhoods by income and crime with centroids",
    x = "Median Income",
    y = "Total Crimes",
    color = "Cluster"
  ) +
  theme_minimal()
```
the black triangles represent the average values of income and crime for each cluster. this shows that crime counts decrease as median income increases. 