---
title: "blog post 5 "
author: "Sean Fung, Zihao Guo, Yawen Zhang"
date: "2024-11-10"
date-modified: "2024-11-10"
draft: FALSE
---

```{r Getting Geographical Data}
#| echo: false
library(tidyverse)
library(sf)
library(tidycensus)
library(here)
library(viridis)
# Load the crime data and convert to spatial format
data <- readRDS(here("filtered_data_2.rds"))
crime_sf <- st_as_sf(data, coords = c("LON", "LAT"), crs = 4326)

# Load neighborhood boundaries
neighborhoods <- st_read(here("LA_Times_Neighborhood_Boundaries-shp/8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp"))
neighborhoods <- st_transform(neighborhoods, st_crs(crime_sf))

# Perform the spatial join to link crime incidents with neighborhoods
crime_with_neighborhoods <- st_join(crime_sf, neighborhoods, join = st_within)

# Load and join median income data
economic_data <- get_acs(
    geography = "block group",
    variables = c(median_income = "B19013_001"),
    state = "CA",
    county = "Los Angeles",
    year = 2020,
    geometry = TRUE,
    cache_table = TRUE
)
economic_data <- st_transform(economic_data, st_crs(crime_with_neighborhoods))
crime_with_income <- st_join(crime_with_neighborhoods, economic_data, join = st_within)

# Summarize crime data by neighborhood, with average income and ensuring valid geometry
crime_summary <- crime_with_income %>%
  group_by(AREA.NAME) %>%
  summarize(
    total_crimes = n(),
    avg_income = mean(estimate, na.rm = TRUE),
    geometry = st_union(geometry)
  ) %>%
  st_as_sf()

# Filter out rows with NA values in avg_income
crime_summary <- crime_summary %>% filter(!is.na(avg_income))
```

```{r Pulling Income Data}


# Pull median income data for Los Angeles
income_data <- get_acs(
  geography = "tract",
  variables = c(median_income = "B19013_001"),
  state = "CA",
  county = "Los Angeles",
  year = 2020,
  geometry = TRUE
)

# Prepare the data for mapping (ensure CRS consistency)
income_data <- st_transform(income_data, crs = st_crs(crime_summary))

# Join income data with crime data (assuming `crime_summary` has neighborhood geometries)
crime_income_map <- st_join(crime_summary, income_data, join = st_within)

# Filter out areas with missing income data (if needed)
crime_income_map <- crime_income_map %>% filter(!is.na(estimate))
```

```{r Income Heatmap}
# Ensure geometries are valid and filter out empty or invalid geometries
# Filter out geometries with coordinates at (0, 0) or zero degrees west (longitude = 0)
crime_income_map <- crime_income_map %>%
  filter(!st_is_empty(geometry)) %>%  # Ensure geometry is not empty
  filter(!sapply(geometry, function(geom) {
    coords <- st_coordinates(geom)
    return(any(coords[,1] == 0 & coords[,2] == 0) || any(coords[,1] == 0) || any(coords[,2] == 0))
  }))

ggplot() +
  geom_sf(data = income_data, aes(fill = estimate), color = NA) +  # Median income layer
  geom_sf(data = crime_summary, aes(color = avg_income), size = 2, alpha = 0.7) +  # Crime data layer
  scale_fill_viridis_c(option = "magma", na.value = "grey50") +
  scale_color_viridis_c(option = "inferno", na.value = "grey50") +
  labs(
    title = "Median Income and Crime Data by Neighborhood in Los Angeles",
    fill = "Median Income",
    color = "Average Income"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  coord_sf(
    xlim = c(-119, -117),  # Adjust these values to focus on the Los Angeles area
    ylim = c(33, 35)       # Adjust these values to focus on the Los Angeles area
  )



```
# crime rates by income levels
Based on the detailed spatial dataset, we're allowed to do some explorations of the relationship between income levels and crime incidents across Los Angeles neighborhoods. 
I'm interested in crime rates by income levels. The chunk below is to determine if crime rates are higher in low-income areas. We categorize neighborhoods into income brackets(low, middle, high income) and calculate crime rates for each bracket. 
```{r}
library(dplyr)

crime_income_map <- crime_summary %>%
  mutate(
    income_bracket = case_when(
      avg_income <= quantile(avg_income, 0.33, na.rm = TRUE) ~ "Low",
      avg_income > quantile(avg_income, 0.33, na.rm = TRUE) & avg_income <= quantile(avg_income, 0.66, na.rm = TRUE) ~ "Middle",
      avg_income > quantile(avg_income, 0.66, na.rm = TRUE) ~ "High"
    )
  )

crime_rate_by_income <- crime_income_map %>%
  group_by(income_bracket) %>%
  summarize(
    avg_crime_rate = mean(total_crimes, na.rm = TRUE)
  )

ggplot(crime_rate_by_income, aes(x = income_bracket, y = avg_crime_rate, fill = income_bracket)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Crime Rate by Income Level",
    x = "Income Bracket",
    y = "Average Crime Rate"
  ) +
  theme_minimal()


```
Observing the graph we get above, it seems that neighborhoods with low-income levels have a higher avergae crime rate compared to middle and high-income areas. I think this trend aligns with common sociological patterns. Now we visualize the distribution of crime rates and income levels in Los Angeles. 
# map of crime rates across income levels
```{r}
library(ggplot2)
library(dplyr)
library(sf)
crime_income_expanded <- crime_income_map %>%
  rowwise() %>%
  mutate(points = list(rep(geometry, total_crimes))) %>%
  unnest(points) %>%
  st_as_sf()

ggplot() +
  geom_sf(data = crime_income_expanded, aes(geometry = points, color = income_bracket), size = 0.1, alpha = 0.6) +
  scale_color_manual(values = c("blue", "purple", "red"), name = "Income Bracket") +
  labs(
    title = "Dot-Density Map of Crime Rates Across Income Levels",
    subtitle = "Each dot represents a crime incident within an income bracket"
  ) +
  theme_minimal() +
  coord_sf(xlim = c(-118.8, -118.1), ylim = c(33.5, 34.5))

```
There's a dense cluster of red/ purple/ blue dots in certain areas, showing that neighborhoods classified under high/ middle/ low income have noticeable crime incidents concentrated within specific locations. 
the visual representation shows that crime rates are not uniformly distributed across all income levels. 
# Input LA Temp CSV and change it to RDS file
```{r}
csvfile <- read.csv("latemp.csv")
saveRDS(csvfile, "latemp.rds")
```
## We had combine the temp data and data that we are using by using left join
```{r}
latemp <- readRDS(here("latemp.rds"))
latemp <- latemp %>%
  mutate(DATE = as.Date(DATE)) %>%
  rename(DATE.OCC = DATE)
latemp_unique <- latemp %>%
  distinct(DATE.OCC, .keep_all = TRUE)
data_with_temp <- data %>%
  left_join(latemp_unique, by = "DATE.OCC")
```
## Find the top 20 temperture that mostly to occur crime
```{r}
top_20_temps <- data_with_temp %>%
  count(TAVG, sort = TRUE) %>%
  top_n(20, n)

top_20_temps
```

## ggplot for temp vs crime occurs
```{r}
ggplot(data_with_temp, aes(x = TAVG)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Crime Count by Temperature",
       x = "Average Temperature (TAVG)",
       y = "Crime Count") +
  theme_minimal()
```
The histogram shows that crime counts are higher in the temperature range of approximately 55°F to 75°F. Within this range, the frequency of crimes appears to be fairly consistent, with a peak around 60°F to 70°F. As temperatures exceed 75°F, there is a noticeable decline in crime counts. This suggests that extremely warm days may be associated with lower crime activity. Similarly, at lower temperatures (below 50°F), there are also fewer crime incidents. This could be due to people staying indoors more on colder days, leading to less social interaction and possibly fewer opportunities for crime.