---
title: "blog post 4"
author: ""
date: "2024-10-31"
date-modified: "2024-10-31"
draft: FALSE
---
```{r}
data <- readRDS("filtered_data_2.rds")

# Save it as a CSV file
write.csv(data, "your_file.csv", row.names = FALSE)
```
Used this to view data in excel file for easier visibility and filtering. 

```{r}
library(ggplot2)
library(here)
data <- readRDS(here("filtered_data_2.rds"))

filtered_data <- subset(data, Descent_Description != "Unknown" & Descent_Description != "Other")

# Create a contingency table for the filtered data
crime_table_filtered <- table(filtered_data$Descent_Description, filtered_data$Crm.Cd.Desc)

chi_test_filtered <- chisq.test(crime_table_filtered)
print(chi_test_filtered)
# Extract standardized residuals
standardized_residuals_filtered <- chi_test_filtered$stdres

top_crimes <- names(sort(colSums(crime_table_filtered), decreasing = TRUE))[1:20]

# Filter the contingency table to only include top crime types
crime_table_top <- crime_table_filtered[, top_crimes]

chi_test_top <- chisq.test(crime_table_top)
standardized_residuals_top <- chi_test_top$stdres


residuals_top_df <- as.data.frame(as.table(standardized_residuals_top))
colnames(residuals_top_df) <- c("Descent", "Crime_Type", "Residual")



ggplot(residuals_top_df, aes(x = Crime_Type, y = Descent, fill = Residual)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "Standardized Residuals of Descent and Top Crime Types",
       x = "Crime Type", y = "Descent Category", fill = "Residual") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

I wanted to ask the question of whether every crime was correlated to descent. Therefore, I conducted a chi-squared test and wanted to see how each race's residual was in comparison to the top 10 crimes (all crimes would be too much). 
- We can see that "White" had a high residual in the burglary category and higher residuals in some other categories as well. We also see lower than average residuals in the aggrevated assault and simple assault categories. 
- As for Hispanic/Mexican, even though they represent such a high amount of crimes (as seen in last blog post), they are moderately around what is expected with the exception of the "Burglary" and "Theft (under $950) categories.
- Other Asian slightly higher in "Burglary"
- "Black" lower in "theft from motor vehicle". 
- "Filipino" higher in "theft from motor vehicle". 
- "Chinese" higher in "theft - grand except guns or livestock". 
- Everything else is mostly the same hovering around the expected with small fluctuations.







